---
title: "Homicides in Europe - Group 11"
output:
  html_document: default
  pdf_document: default
date: "2024-03-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TASK A

kings \<- read_csv("data/Danske_konger_Gruppe11.csv") names(kings)

kings %\>% mutate(duration = `Slut paa regerings tid` - `Start paa regerings tid`) %\>% mutate(midyear = `Start paa regerings tid` - duration/2) %\>% ggplot(aes(x = midyear, y = duration)) + geom_point()+ geom_smooth()

```{}
```

# Task B

## Loading data

```{r}
library(tidyverse)
```

```{r}
Western_Europe <- read_csv("data/homicide-rates-across-western-europe.csv")
```

Checking the data:

```{r}
head(Western_Europe)
```

Renaming "Homicide rate in Europe over long-term (per 100,000)" to "homicides_per_100k"

```{r}
names(Western_Europe)[4] <- "homicides_per_100k"
```

## Seeing the long-term trend in homicides

```{r}
Western_Europe %>% 
  ggplot(aes(x = Year, y = homicides_per_100k, color = Entity)) + 
  geom_line()
```

```{r}
eval(Western_Europe)
```

## Making it easier to see

```{r}
Western_Europe %>% 
  ggplot(aes(x = Year, y = homicides_per_100k, color = Entity)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(title="Homicides in Europe",
       x = "Year", 
       y = "homicides_per_100k") + 
  facet_wrap(~ Entity)
```

## Now we have a look at our Danish Rulers Data Set

```{r}
library(readr)
Danske_konger_Gruppe11 <- read_csv("data/Danske_konger_Gruppe11.csv")
```

## Plotting the data

```{r}
Danske_konger_Gruppe11 %>% 
  mutate(duration = `Slut paa regerings tid` - `Start paa regerings tid`) %>% 
  mutate(midyear = `Start paa regerings tid` - duration/2) %>% 
  ggplot(aes(x = midyear, y = duration)) + 
  geom_smooth()
```
